\section{Conclusion}
\label{conclusion}

We introduced Agent Bazaar, a multi-agent marketplace simulator that exposes two systemic failure modes of unaligned LLM agents: the \textbf{B2C Crash}, where myopic undercutting drives prices below unit cost, and the \textbf{C2C Lemon Market}, where a Sybil Principal exploits cheap LLM identity generation to flood the market with fraudulent listings.
Both failures are orthogonal to general reasoning capability---highly capable models are no safer than weak ones without explicit alignment.
Our Economic Alignment Finetuning procedure (SFT on EAS-filtered market traces via LoRA) produces \textbf{Stabilizing Firms} that hold prices above the competitive floor and \textbf{Skeptical Guardians} that reject Sybil clusters with high accuracy, restoring Nash equilibrium in both settings.

\textbf{Limitations.}
Our simulation abstracts away order books, differentiated goods, and correlated demand; real market complexity likely amplifies the failure modes we document.
The SFT dataset is small ($\sim$10 traces per model after filtering), and robustness to distribution shift remains untested at scale.
Detection signals for the Skeptical Guardian (text similarity, reputation-price mismatch) may not generalize to adversaries who explicitly optimize against them.

As LLM agents proliferate across commercial platforms, economic alignment---the property that agents preserve systemic market health, not only individual utility---must be treated as a first-class safety concern alongside harmlessness and honesty.
Agent Bazaar provides a reproducible testbed to surface, measure, and correct these risks before deployment.
