\section{Experiments}
\label{exp}

We conduct three sets of experiments to demonstrate the fragility of base models and the efficacy of our alignment finetuning.

\subsection{Exp 1: The Crash (B2C Instability)}
\textbf{Setup:} We instantiate Environment A (Section~\ref{setup:crash}) with $N=5$ firm agents endowed with \$500 in cash and $M=50$ consumer agents. Each episode runs for $T=365$ timesteps. Consumer demand is drawn from a Poisson process $D_t \sim \mathrm{Poisson}(\lambda=30)$ per timestep. The token budget is $\tau = 3$, meaning each firm observes at most 3 competitor prices per round.
\textbf{Baseline Result:} Gemini-2.5 flash base agents aggressively undercut prices, causing price collapse below unit costs until one agent remains in business. The last remaining firm exploits their monopolistic position to maximize revenue. 
\textbf{Untrained Intervention:} Stabilizing firm agents coupled with limited discovery for consumers decreases collapse probability. 
\textbf{Trained Intervention:} FIRMS FINETUNED ON "SURVIVAL TRACES" (TOP 10\% OF EPISODES) LEARN TO SIGNAL AND MAINTAIN A PRICE FLOOR, REDUCING BANKRUPTCY RATE TO 10\%.


\subsection{Exp 2: The Lemon Market (C2C Fraud)}
\textbf{Setup:} 10 Sellers (5 Honest, 5 Deceptive), 10 Buyers. 
\textbf{Result:} Base Buyers trust descriptions $D$ blindly. Deceptive sellers capture 90\% of revenue initially, but Buyer Welfare becomes negative. Market volume collapses by $t=20$.
\textbf{Intervention:} "Guardian" Buyers finetuned to detect discrepancies between $D$ and $R$ successfully boycott deceptive sellers. Deceptive revenue drops to 10\%, forcing a strategy shift or exit.

\subsection{Exp 3: Systemic Resilience (Shocks)}
\textbf{Setup:} We introduce a "Supply Shock" (Cost doubles) in the Amazon scenario and a "Flood of Fakes" (90\% Deceptive) in the eBay scenario.
\textbf{Result:} Aligned agents recover equilibrium 4x faster than base agents.

\subsection{Model Family Benchmarking}
We compare the performance of different open-source model families in the marketplace environment. Our study includes:
\begin{itemize}
    \item \textbf{Gemma 3}: Google's latest open model, optimizing efficient reasoning at the 4B/8B scale.
    \item \textbf{Qwen 3}: Alibaba's frontier model, known for superior tool-use and structured output adherence.
    \item \textbf{Ministral 3}: Mistral AI's edge-optimized model, exhibiting high reasoning density for its size.
    \item \textbf{OLMo 3}: The fully open scientific model from AI2, ensuring results generalize to non-proprietary training data.
    \item \textbf{Llama 3.2}: Meta's multimodal-ready model, serving as a robust baseline for general purpose agents.
\end{itemize}
We analyze the Pareto frontier of \textit{Compute Cost} vs. \textit{Economic Stability}, demonstrating that modern 3rd-generation open weights models achieve significant survival advantages over previous generations when finetuned on market dynamics.
