\documentclass{article}
\usepackage{colm2026_conference}

\usepackage{microtype}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

% \title{The Agent Bazaar: Benchmarking Economic Alignment in High-Frequency Multi-Agent Ecosystems}
\title{Agent Bazaar: Enabling Economic Alignment in Multi-Agent Marketplaces}

\author{Seth Karten \\
Princeton University \\
\texttt{sethkarten@princeton.edu} \\
\And
Cameron Crow \\
Princeton University \\
\texttt{cameroncrow@princeton.edu} \\
\And
Chi Jin \\
Princeton University \\
\texttt{chijin@princeton.edu}
}

\begin{document}

\maketitle

\begin{abstract}
The deployment of Large Language Models (LLMs) as autonomous economic agents introduces systemic risks that extend beyond individual capability failures. As agents transition to directly interacting with marketplaces (e.g., Amazon, eBay), their collective behavior can amplify volatility and mask deception at scale. We introduce \textbf{Agent Bazaar}, a multi-agent simulation framework for evaluating \textit{Economic Alignment}, defined as the capacity of agentic systems to preserve market stability and integrity. Critically, economic alignment is distinct from general reasoning capability: a highly capable agent can still be economically unsafe. We identify two emergent failure modes: (1) \textit{Algorithmic Instability} in the scenario of a business-to-consumer (B2C) market, ``The Crash'', where myopic firms amplify price volatility until the market collapses, and (2) \textit{Sybil Deception} in the scenario of a consumer-to-consumer market (C2C), ``The Lemon Market'', where a single deceptive principal operating multiple coordinated seller identities floods the market with fraudulent listings, eroding trust, trading volume, and consumer welfare. First, we evaluate the economic alignment score (EAS) of frontier and open-weight models across a variety of model sizes in each scenario. We demonstrate that without intervention, models largely fail to self-regulate in both scenarios, leading to market collapse in The Crash and exploitation in The Lemon Market. To address this, we propose an alignment harness: using \textit{Stabilizing Firms} to act as volatility dampeners, and \textit{Skeptical Guardians} to detect Sybil coordination patterns in seller networks. Even with these agents, almost all models (except for Claude Sonnet 4.6 in The Crash) fail to self-regulate in both scenarios, with Claude barely achieving stability in The Crash. We further show that with RL finetuning of open-weights models, economic alignment can be achieved in both scenarios.
% The deployment of Large Language Models (LLMs) as autonomous economic agents introduces systemic risks that extend beyond individual capability failures. As agents transition to high-frequency marketplaces (e.g., Amazon, Moltbook) executing over one million agent-driven transactions daily, their collective behavior can amplify volatility and mask deception at scale. We introduce \textbf{Agent Bazaar}, a multi-agent simulation framework for evaluating \textit{Economic Alignment}---defined as the capacity of agentic systems to preserve market stability and integrity. Critically, economic alignment is distinct from general reasoning capability: a highly capable agent can still be economically unsafe. We identify two emergent failure modes: (1) \textit{Algorithmic Instability} (``The Crash''), where myopic firms amplify price volatility in a business-to-consumer (B2C) market until it collapses, and (2) \textit{Sybil Deception} (``The Lemon Market''), where a single deceptive principal operating multiple coordinated seller identities floods a consumer-to-consumer (C2C) market with fraudulent listings, eroding trust, trading volume, and consumer welfare. To address this, we propose a dual-alignment strategy: \textit{Stabilizing Firms} to act as volatility dampeners, and \textit{Skeptical Guardians} to detect Sybil coordination patterns in seller networks. We evaluate this strategy via prompt-conditioned variants across frontier models (Gemini 3 Flash, GPT 5.4, Claude Sonnet 4.6) and supervised finetuning of Qwen 3.5 9B. We demonstrate that without intervention, models largely fail to self-regulate under both failure modes, leading to market collapse. Among the frontier models, only Claude Sonnet 4.6 achieved stability in The Crash without stabilizing firm prompt-conditioning; Gemini 3 Flash and GPT 5.4, with and without intervention, experienced bankruptcy rates of 80-100\% and monopolistic exploitation. In the Lemon Market, unconditioned buyers demonstrate an improved ability to detect fraudulent sellers when such identities occupy a larger portion of the seller space, but suffer lower welfare and allow deceptive sellers to steal increasing revenue shares. We evaluate Qwen 3.5 in both base and finetuned configurations. our results show that aligned agents do not merely maximize profit; they maintain market stability, protecting consumer welfare and trading volumes in non-stationary environments. These findings indicate that economic alignment can be framed as a design target for autonomous economic agents to limit the risks of consumer exploitation and market collapse in B2C and C2C environments. 
\end{abstract}
\input{sections/introduction}
\input{sections/related_work}
\input{sections/problem_setup}
\input{sections/methodology}
\input{sections/experimental_setup}
\input{sections/results}
\input{sections/conclusion}

\bibliographystyle{colm2026_conference}
\bibliography{references}

\input{sections/appendix}



\end{document}
